No proof nets for MLL with units\\[5pt]\Large Proof equivalence in MLL is \textsc{pspace}-complete

### 2014-01-14 main/LICS.tex
first stab at intro

\usepackage{amssymb,amsthm,amsmath}
\usepackage{microtype}
%\usepackage{enumitem}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{mathtools}
\usepackage[round]{natbib}
\usepackage{mllequiv,willemtools}

@@

\makeTheoremDefs

\author{
  \IEEEauthorblockN{Willem Heijltjes}
  \IEEEauthorblockA{
  		University of Bath \\
		Claverton Down \\
		Bath BA2~7AY\\
		w.b.heijltjes@bath.ac.uk
	}
\and
  \IEEEauthorblockN{Robin Houston}
  \IEEEauthorblockA{Kiln Enterprises}
}

\title{The proof equivalence problem for multiplicative linear logic is \textsc{pspace}-complete}

@@

\begin{abstract}
MLL proof equivalence is the problem of deciding whether two proofs in multiplicative linear logic are related by a series of inference permutations.
%
It is also known as the word problem for $*$-autonomous categories.
%
Previous work has shown the problem to be equivalent to a rewiring problem on proof nets, which are not canonical for full MLL due to the presence of the two units.
%
Drawing from recent work on reconfiguration problems, in this paper it is shown that MLL proof equivalence is PSPACE-complete, using a reduction from Nondeterministic Constraint Logic.
%
The result rules out the existence of a satisfactory notion of proof nets for MLL with units (under current complexity assumptions).
\end{abstract}

@@

\begin{figure}
%%\noindent
%\fbox{\MLLrule b}
%\hfill\fbox{\MLLrule 1}
%\hfill\fbox{\MLLrule p}
%\hfill\fbox{\MLLrule t}
%\hfil
\[
\begin{array}{llll}
	\MLLrule b
 &	\MLLrule 1
 &	\MLLrule p
 &	\MLLrule t
\end{array}
\]
\caption{Inference rules for unit-only \MLL}
\label{fig:MLL}
\end{figure}

\begin{figure}
\[
	\vc{\MLLperm{bb1}} \perm \vc{\MLLperm{bb2}}
\qquad
	\vc{\MLLperm{bp1}} \perm \vc{\MLLperm{bp2}}
\]
\[
	\vc{\MLLperm{bt1}} \perm \vc{\MLLperm{bt2}} \perm \vc{\MLLperm{bt3}}
\]
\[
	\vc{\MLLperm{pp1}} \perm \vc{\MLLperm{pp2}}
\]
\[
	\vc{\MLLperm{pt1}} \perm \vc{\MLLperm{pt2}}
\]
\[
	\vc{\MLLperm{tt1}} \perm \vc{\MLLperm{tt2}}
\]
\caption{Permutations}
\label{fig:permutations}
\end{figure}

@@

@@

%------------------------------------------------------------------------------------------------

@@

\bibliographystyle{plainnat}
\bibliography{refs}

>>>>

\usepackage{amssymb,amsthm,amsmath}
\usepackage{microtype}
%\usepackage{enumitem}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{mathtools}

% Sorts citations; e.g. [1], [2], [5]--[7]
%\usepackage{cite}

% More reference handling options
%\usepackage{natbib}\renewcommand\bibfont{\footnotesize}

\usepackage{mllequiv,willemtools}

@@

\shortJournalNames
\makeTheoremDefs

\author{
  \IEEEauthorblockN{Willem Heijltjes}
  \IEEEauthorblockA{
  		University of Bath
	\\	Claverton Down
	\\	Bath BA2~7AY
	\\	w.b.heijltjes@bath.ac.uk
	}
\and
  \IEEEauthorblockN{Robin Houston}
  \IEEEauthorblockA{
		Kiln Enterprises
	\\	robin.houston@gmail.com	
	}
}

%\title{The proof equivalence problem for multiplicative linear logic is \textsc{pspace}-complete}

%\title{Proof equivalence in MLL is \textsc{pspace}-complete}

\title{No proof nets for MLL with units\\[5pt]\Large Proof equivalence in MLL is \textsc{pspace}-complete}

%============================================================

@@

\begin{abstract}
MLL proof equivalence is the problem of deciding whether two proofs in multiplicative linear logic are related by a series of inference permutations.
%
It is also known as the word problem for $*$-autonomous categories.
%
Previous work has shown the problem to be equivalent to a rewiring problem on proof nets, which are not canonical for full MLL due to the presence of the two units.
%
Drawing from recent work on reconfiguration problems, in this paper it is shown that MLL proof equivalence is PSPACE-complete, using a reduction from Nondeterministic Constraint Logic.
%
An important consequence of the result is that the existence of a satisfactory notion of proof nets for MLL with units is ruled out (under current complexity assumptions).
\end{abstract}

%------------------------------------------------------------

\section{Introduction}

An important element of the theory of linear logic \cite{Girard-1987} is the idea that proofs should be considered modulo the permutation of inferences.
%; that the chosen order of permutable inferences is inessential, or `bureaucracy'.
%contributes no essential proof content, and should be considered `bureaucracy'.
%
%The idea is supported by the categorical semantics of linear logic, in which proofs up to permutations denote the same morphism. 
%
\emph{Proof equivalence} is the problem of deciding whether two proofs are equal modulo permutations.

The proof equivalence problem is closely related to the notion of \emph{proof nets}.
%
These are canonical, geometric representations of proof that factor out permutations, so that proofs are equivalent if and only if they have the same underlying proof net.
%
Canonical proof nets have been found for several fragments of linear logic: the multipicative fragment without units \cite{Girard-1987,Danos-Regnier-1989}, the combined multiplicative-additive fragment without units \cite{Hughes-vanGlabbeek-2005}, and the additive fragment, including the additive units \cite{Heijltjes-2011}.
%
For the full multiplicative fragment, with units, several notions of proof net have been proposed \cite{BCST,Strassburger-lamarche-2004,HughesFreeStar}.
%
However, none is canonical, and whether a satisfactory notion of proof net can be found has remained an open question.

In this paper we establish that the proof equivalence problem for multiplicative linear logic with units is \textsc{pspace}-complete.
%
As a main corrolary, this effectively rules out the possibility of finding a satisfactory notion of proof net for this fragment.

%------------------------------------------------------------

@@

\begin{figure}[!b]
\quad\MLLrule b
\hfill\MLLrule 1
\hfill\MLLrule p
\hfill\MLLrule t
\quad
\caption{Inference rules for unit-only \MLL}
\label{fig:MLL}
\end{figure}

\begin{figure}[!tb]
\renewcommand\scalefactor{0.88}
\[
\begin{array}{@{}rcl@{}}
	\multicolumn{3}{@{}c@{}}{
		\vc{\scale{\MLLperm{bb1}}} ~\perm~ \vc{\scale{\MLLperm{bb2}}}
		\hfill
		\vc{\scale{\MLLperm{bp1}}} ~\perm~ \vc{\scale{\MLLperm{bp2}}}
	}
\\ \\[6pt]
	\multicolumn{3}{@{}c@{}}{
		\vc{\scale{\MLLperm{bt2}}} ~\perm~ \vc{\scale{\MLLperm{bt1}}} ~\perm~ \vc{\scale{\MLLperm{bt3}}}
	}
\\ \\[6pt]
	\vc{\scale{\MLLperm{pp1}}} &\perm& \vc{\scale{\MLLperm{pp2}}}
\\ \\[6pt]
	\vc{\scale{\MLLperm{pt1}}} &\perm& \vc{\scale{\MLLperm{pt2}}}
\\ \\[6pt]
	\vc{\scale{\MLLperm{tt1}}} &\perm& \vc{\scale{\MLLperm{tt2}}}
\end{array}
\]
\caption{Permutations}
\label{fig:permutations}
\end{figure}

%===========================
\bibliographystyle{plain}
\bibliography{../refs}
\end{document}
%===========================

@@

%------------------------------------------------------------

@@

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

@@

%------------------------------------------------------------

\section{Conclusions}
